State of the Practice for Lattice Boltzmann Method Software

### 2022-01-27 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Updated section 3.7 on Maintainability, incorporated information from Peter #121

The software packages were tested for handling unexpected input, including
incorrect data types, empty input, and missing files or links. Success
predicated on a reasonable response from the system, including appropriate error
messages and an absence of unrecoverable system failures.

@@

Figure~\ref{Fig_Robustness} shows the surface robustness ranking of the software
packages using AHP. Software packages with a high score behaved reasonably in
response to unexpected input as described above. All of the software packages
that installed correctly passed this test. They output descriptive error
messages or did not crash. Software packages with a lower surface robustness
score did not install correctly, so their robustness score may not be a true
reflection of runtime robustness. All successfully installed software packages
that require a plain text input file correctly handled an unexpected change to
these input files, including a replacement of new lines with carriage returns.
Furthermore, nine of the top 10 ranked packages are noted as being alive. LIMBES
is noted as being classified as dead.

@@

Although the software packages all apply LBM to solve scientific computing
problems, the packages focus on varied CFD problems, with varying parameters,
and are technically different from each other (as shown in
Table~\ref{tbl_features}). Due to this, a comparison of performance is not
appropriate. In this project we instead looked through each software package's
artifacts for evidence that performance was considered. The artifacts of 18
software packages mentioned parallelization. This included GPU processing and
the CUDA parallel computing platform, which were mentioned in the artifacts of 6
packages (ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB). GPUs provide
superior processing power and speed compared to CPUs, and are often used for
scientific computing when a large amount of data is involved. The software
package TCLB is implemented in a highly efficient multi-GPU code to achieve
performance suitable for model optimization \citep{rutkowski2020open}. In the
Ludwig package, a so-called mixed mode approach is used where fine-grained
parallelism is implemented on the GPU, and MPI is used for even larger scale
parallelism \citep{gray2013ludwig}. While one software package (Sailfish)
required CUDA and GPU processing, some (ESPResSo, lbmpy, lettuce, pyLBM, TCLB)
have the option of using either the GPU or the CPU. The packages that require
GPU and CUDA have better performance at the expense of installability and
surface reliability.

@@

Software package artifacts were reviewed for the presence of a tutorial, a user
manual, documented user characteristics, and a user support model. In total 19
software packages have a tutorial, 13 have a user manual, and 11 have both. The
tutorials vary in scope and substance, and eight include expected output. Most
user manuals are in the form of a file that can be downloaded, while some are
rendered on a web-page. Some packages (waLBerla) do not have a user manual, but
do have useful documentation distributed on their web-pages. Expected user
characteristics are documented in five software packages (laboetie, LIMBES,
Ludwig, Musubi, Palabos). Users are typically scientists or engineers. Their
background is often physics, chemistry, biophysics, or mathematics. All but one
of the packages (LIMBES) have a user support model, and many of them have
multiple avenues of user support. The most popular avenue of support is Git,
followed by email and forums. One software package (OpenLB) has an FAQ page.

@@

Figure~\ref{Fig_Usability} shows the surface usability ranking of the software
packages using AHP. Software packages with a high score have a tutorial and user
manual, sometimes have documented user characteristics, and have at least one
user support model. Many packages have several user support models. Furthermore,
four of the top five ranked packages are noted as being alive.

@@

Issue tracking is used in 23 software packages, 15 of which use Git, seven use
email, and one (SunlightLB) uses SourceForge. Most software packages that use
Git have most of their issues closed, and only three (laboetie, lettuce,
Sailfish) have less than 50 percent of their issues closed. All of the top five
overall ranked packages have most of their issues closed. Alive packages (11 use
Git issue tracking) have 64\% of their issues closed, while dead packages (3 use
Git issue tracking) have 71\% of their issues closed. This information is
presented in Table~\ref{gitrepodata}. Furthermore, 13 packages that use Git for
issue tracking use GitHub as a version control system, while two (Palabos,
waLBerla) use GitLab. Of the other packages, one package (SunlightLB) uses CVS
for issue tracking, and seven packages do not appear to use any issue tracking
system.

\begin{table}
	\begin{center}
		\begin{tabular}{ p{3.5cm}p{3.5cm}p{3.5cm}p{2.5cm} }
			\toprule
			Name & $\%$ Issues Closed & $\%$ Code Comments & Status\\
			\midrule
			MP-LABS & 100.00 & 26.67 & Dead\\
			Musubi & Not Git & 24.19 & Alive\\
			waLBerla & 72.90 & 22.62 & Alive\\
			OpenLB & Not Git & 22.43 & Alive\\
			ESPResSo & 89.26 & 21.78 & Alive\\
			Ludwig& 60.00 & 20.70 & Alive\\
			Palabos & 89.47 & 17.76 & Alive\\
			SunlightLB & Not Git & 17.67 & Dead\\
			LIMBES & Not Git & 17.39 & Dead\\
			ESPResSo++ & 66.28 & 17.10 & Alive\\
			HemeLB & No Issues & 16.68 & Dead\\
			pyLBM & 66.67& 16.12 & Alive\\
			MechSys & Not Git & 15.11 & Alive\\
			LB3D-Prime & Not Git & 14.34 & Dead\\
			LB3D & Not Git & 13.76 & Dead\\
			LB2D-Prime & Not Git & 13.61 & Dead\\
			Sailfish & 22.22 & 9.26 & Alive\\
			lettuce & 33.33 & 8.19 & Alive\\
			DL\_MESO (LBE) & Not Git & 8.06& Alive\\
			TCLB & 60.32 & 6.02 & Alive\\
			laboetie & 18.75 & 2.47 & Dead\\	
			lbmpy& 58.33  & 2.03 & Alive\\	
			LatBo.jl & 93.33 & 0.40 & Dead\\
			LUMA& 85.71   & 0.20 & Alive\\
			\bottomrule
		\end{tabular}
		\caption{Git Repository Data} \label{gitrepodata}
	\end{center}
	\end{table}

Software package code files were further measured for the percentage of code
that is comments. The findings are presented in Table~\ref{gitrepodata}.
Packages with a higher percentage of comments were designated as more
maintainable. Comments represent more than 10 percent of code files in 16
packages, and the average percentage of code comments is about 14 percent. All
of the top five overall ranked packages have more than the average. LUMA has only 0.2 percent comments, the fewest of any package. This package has the most lines of source code, with over four million. The next largest package is ESPResSo++ with one million.

@@

Figure~\ref{Fig_Maintainability} shows the maintainability ranking of the
software packages using AHP. Software packages with a high score provide version
numbers on documents and source code releases, have an abundance of high quality
artifacts, and use an issue tracking tool and version control system. These
packages also appear to reasonably handle issue tracking, having most of their
issues closed. Their code files are well commented with more than 10 percent of
the code being comments. Furthermore, four of the top 5 ranked packages are
noted as being alive. MP-LABS is noted as being dead.\\

\wss{Sort ``Git Repository Data'' table in descending order of percentage of
code comments.}

>>>>

The software packages were tested for handling unexpected input, including
incorrect data types, empty input, and missing files or links. Success is
predicated on a reasonable response from the system, including appropriate error
messages and an absence of unrecoverable system failures.

@@

Figure~\ref{Fig_Robustness} shows the surface robustness ranking of the software
packages using AHP. Software packages with a high score behaved reasonably in
response to unexpected input as described above. All of the software packages
that installed correctly passed this test. They output descriptive error
messages or at least do not crash. Software packages with a lower surface
robustness score did not install correctly, so their robustness score may not be
a true reflection of runtime robustness. All successfully installed software
packages that require a plain text input file correctly handled an unexpected
change to these input files, including a replacement of new lines with carriage
returns. Furthermore, nine of the top 10 ranked packages are noted as being
alive.

@@

Although the software packages all apply LBM to solve scientific computing
problems, the packages focus on varied CFD problems, with varying parameters,
and are technically different from each other (as shown in
Table~\ref{tbl_features}). Due to this, a comparison of run-time performance is
not appropriate. Instead we looked through each software package's artifacts for
evidence that performance was considered. The artifacts of 18 software packages
mentioned parallelization. This included GPU processing and the CUDA parallel
computing platform, which were mentioned in the artifacts of 6 packages
(ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB). When a high degree of
parallelization is possible GPUs provide superior speed compared to CPUs. The
software package TCLB is implemented in a highly efficient multi-GPU code to
achieve performance suitable for model optimization \citep{rutkowski2020open}.
In the Ludwig package, a so-called mixed mode approach is used where
fine-grained parallelism is implemented on the GPU, and MPI is used for even
larger scale parallelism \citep{gray2013ludwig}. While one software package
(Sailfish) required CUDA and GPU processing, some (ESPResSo, lbmpy, lettuce,
pyLBM, TCLB) have the option of using either the GPU or the CPU. In general, the
packages that require GPU and CUDA have better performance at the expense of
installability and surface reliability.

@@

Software package artifacts were reviewed for the presence of a tutorial, a user
manual, documented user characteristics, and a user support model. In total 19
software packages have a tutorial, 13 have a user manual, and 11 have both. The
tutorials vary in scope and substance, and eight include expected output. Most
user manuals are in the form of a file that can be downloaded, while some are
rendered on a web-page. Some packages (waLBerla) do not have a user manual, but
do have useful documentation distributed throughout their web-pages. Expected
user characteristics are documented in five software packages (laboetie, LIMBES,
Ludwig, Musubi, Palabos). Users are typically scientists or engineers. Their
background is often physics, chemistry, biophysics, or mathematics. All but one
of the packages (LIMBES) have a user support model, and many of them have
multiple avenues of user support. The most popular avenue of support is Git,
followed by email and forums. One software package (OpenLB) has an FAQ page.

@@

Figure~\ref{Fig_Usability} shows the surface usability ranking of the software
packages using AHP. Software packages with a high score have a tutorial and user
manual, sometimes have documented user characteristics, and have at least one
user support model. Many packages have several user support models. Furthermore,
four of the top five packages are noted as being alive.

@@

Issue tracking is used in 23 software packages, 15 of which use GitHub or GitLab
to host their git repository, seven use email, and one (SunlightLB) uses
SourceForge. Most software packages that use GitHub/GitLab have most of their
issues closed, and only three (laboetie, lettuce, Sailfish) have less than 50
percent of their issues closed. All of the top five overall ranked packages have
most of their issues closed. Alive packages (11 use issue tracking) have 64\% of
their issues closed, while dead packages (3 use issue tracking) have 71\% of
their issues closed. This information is presented in Table~\ref{gitrepodata}.
Furthermore, 13 packages that use git for version control use GitHub, while two
(Palabos, waLBerla) use GitLab. Of the other packages, one package (SunlightLB)
uses CVS for issue tracking, and seven packages do not appear to use any issue
tracking system.

\begin{table}
	\begin{center}
		\begin{tabular}{ p{3.5cm}p{3.5cm}p{3.5cm}p{2.5cm} }
			\toprule
			Name & $\%$ Issues Closed & $\%$ Code Comments & Status\\
			\midrule
			MP-LABS & 100.00 & 26.67 & Dead\\
			Musubi & Not Git & 24.19 & Alive\\
			waLBerla & 72.90 & 22.62 & Alive\\
			OpenLB & Not Git & 22.43 & Alive\\
			ESPResSo & 89.26 & 21.78 & Alive\\
			Ludwig& 60.00 & 20.70 & Alive\\
			Palabos & 89.47 & 17.76 & Alive\\
			SunlightLB & Not Git & 17.67 & Dead\\
			LIMBES & Not Git & 17.39 & Dead\\
			ESPResSo++ & 66.28 & 17.10 & Alive\\
			HemeLB & No Issues & 16.68 & Dead\\
			pyLBM & 66.67& 16.12 & Alive\\
			MechSys & Not Git & 15.11 & Alive\\
			LB3D-Prime & Not Git & 14.34 & Dead\\
			LB3D & Not Git & 13.76 & Dead\\
			LB2D-Prime & Not Git & 13.61 & Dead\\
			Sailfish & 22.22 & 9.26 & Alive\\
			lettuce & 33.33 & 8.19 & Alive\\
			DL\_MESO (LBE) & Not Git & 8.06& Alive\\
			TCLB & 60.32 & 6.02 & Alive\\
			laboetie & 18.75 & 2.47 & Dead\\	
			lbmpy& 58.33  & 2.03 & Alive\\	
			LatBo.jl & 93.33 & 0.40 & Dead\\
			LUMA& 85.71   & 0.20 & Alive\\
			\bottomrule
		\end{tabular}
		\caption{Git Repository Data, Sorted by Percentage of Code that is
		Comments} \label{gitrepodata}
	\end{center}
	\end{table}

Software package code files were further measured for the percentage of code
that is comments. The findings are presented in Table~\ref{gitrepodata}, sorted
by the percentage. Packages with a higher percentage of comments were assumed to be more maintainable. Comments represent more than 10 percent of code files in
16 packages, and the average percentage of code comments is about 14 percent.
All of the top five overall ranked packages have more than the average. LUMA has
only 0.2 percent comments, the fewest of any package. This package has the most
lines of source code, with over four million. The next largest package is
ESPResSo++ with one million.

@@

Figure~\ref{Fig_Maintainability} shows the maintainability ranking of the
software packages. Software packages with a high score provide version numbers
on documents and source code releases, have an abundance of high quality
artifacts, and use an issue tracking tool and version control system. These
packages also appear to reasonably handle issue tracking, having most of their
issues closed. Their code files are well commented with more than 10 percent of
the code being comments. Furthermore, four of the top 5 ranked packages are
noted as being alive.
